Selecting and ranking interesting pairs of cyclic temporal granularities

### 2020-11-05 main.Rmd
bounds deleted

<!-- Scagnostics literature -->
This is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which is used to discern meaningful patterns in large collections of scatterplots. Given a set of $v$ variables, there are $v(v-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots. Therefore
for even small $v$, the number of scatterplots can be
large, and scatterplot matrices (SPLOMs) could easily run out of pixels when presenting high-dimensional data. @Dang2014-tw and @wilkinson2005graph provides potential solutions to this, where few characterizations help us to locate anomalies for further analysis or search for similar distributions in a "large" SPLOM with more than a hundred dimensions.

@@

# Median Maximum Pairwise Distances (MMPD)

## Principle

@@

## Computation

### Distance between distributions

@@

### Normalize distances

@@

### Distribution of distances

### Theoretical evidence

JS distances are distributed as chi-squared with $m$ df where we discretize the continuous distribution with $m$ discrete values. Taking sample percentiles to approximate the integral would mean taking $m = 99$.
With large $m$, chi-squared is asymptotically normal by the CLT. Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution. Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made.

### Empirical evidence

Distribution of JS distances is assumed to be normal but the mean and variance are estimated from the sample, rather than deducing it from the number of discretization used to approximate the continuous distribution. We look at different scenarios, where observations are collected from Normal, Exponential, Chi-squared and Gumbel distribution and found the distribution of JS distances are similar, irrespective of which distribution they are drawn from.

#### Initial distribution of observed variables shown in plot title

@@

## Algorithm

@@

### Bounds

This is not correct because MMPD should be median of standardized Gumbel distribution. So no bound?

By @Lin1991-fj,
$$ 0 \leq JSD(P||Q) \leq ln(2)$$.

Thus,  $$ 0 \leq MMPD \leq \frac{ln(2)}{ln(k)}$$.
Now, by assumption $k \geq 2$ and hence,
<!-- \begin{equation} -->
<!-- For k = 2, $\frac{ln(2)}{ln(k)} = 1$  -->
<!-- \end{equation} -->
<!-- For $k > 2$, $\frac{ln(2)}{ln(k)}  <1$. -->

\[\frac{ln(2)}{ln(k)} : \left\{
  \begin{array}{lr}
    1 & if \ k = 2\\
    <1 & if \ k \ge 2
  \end{array}
\right.
\]

Thus, $$ 0 \leq MMPD \leq \ 1$$

@@

## Design and analysis of simulation experiments

The behavior of the test statistic is monitored and understood through simulation experiments. A single simulation consisting of computational operations on a panel generating MMPD the value of which represents to what extent a pair of cyclic granularities would be interesting when displayed in the design. In comparing different simulation runs, we distinguish between different distributions and simulation scenarios (also iterations may be).

### Simulated designs:

 D1. Same distribution of the measured variable across all x-axis and facet categories
 D2. Different distribution across facet categories but same across x-axis categories
 D3. Different distribution across x-axis categories but same across facet categories
 D4. Different across both x-axis and facet categories

Each of these case scenarios tried against following distribution to check if underlying distribution has a role to play.

 a) Normal 
 b) Exponential
 c) Gamma
 d) Weibull

### Environment

R version 4.0.1 (2020-06-06) is used with platform: x86_64-apple-darwin17.0 (64-bit) running under: macOS Mojave 10.14.6

### Experimental set up

This section desctibes the design of simulation runs, in terms of the scenarios simulated, the number of permutations used to compute estimates and to display the distribution of MMPD.

 <!-- - seed while generating random observations `seed1 = set.seed(9999)`   -->
 <!-- - grid with how many x levels `range_nx_vec = seq(2:4)`   -->
 <!-- - grid with how many facet levels `range_nfacet_vec = seq(2:4)`   -->
 <!-- - number of observations to be generated `ntimes_val = 500`   -->
 <!-- - number of permutation of data done to compute estimates `nperm_val = 2`   -->
 <!-- - number of permutations of data done to display histogram for each panel `nsim_val = 10`    -->
 <!-- - ordered or un-ordered categories of x levels while computing distances `dist_ordered_val = TRUE`   -->
 <!-- - distance between which quantiles are computed `quantile_prob_vec = seq(0.01, 0.99, 0.01)`   -->

```{r}
knitr::include_graphics("Figs/simulation.png")
```

Observations from the null distribution are chosen as:  
`sim_norm_null =   distributional::dist_normal(5, 10)`  
`sim_exp_null  =   distributional::dist_expoential(5)`  
`sim_gamma_null  =   distributional::dist_gamma(0.5, 2)`  
`sim_weibull_null  =   distributional::dist_weibull(1.5, 3)`

The various designs are selected by increasing the mean/rate/shape of these distributions proportionally with that parameter.

<!-- To check if different distributions impact, simulate data from different distributions. A 3 way contingency table with number of facet levels, x-axis levels and distribution type should be constructed with each cell value as the p-value. -->

<!-- Table -->
<!-- Conclusions -->
<!-- (in Null distribution) -->

<!-- #### Varying distribution across facet -->
<!-- #### Varying distribution across x-axis -->
<!-- #### Varying distribution across both facets and x-axis -->
<!-- #### Repeat all with varying facet and x-axis levels -->

<!-- Table -->
<!-- Conclusions -->
<!-- (in power) -->

### Simulation results

#### Null distribution

Under the null hypothesis, all the combinations of x-axis and facets are obtained from the same distribution. We need to study the distribution of the mmpd values under the null hypothesis to see if comparison of their values across different x-axis and facet levels are at all possible.

```{r}
# all_files = list.files(path = "../hakear-drake/data", pattern = ".rds")
# all_files
# 
# read_files = lapply(all_files, function(x){
#   read_rds(paste0("../hakear-drake/data/", x))
# })
# 
# names(read_files) = str_remove(all_files, pattern = ".rds")
# 
# null_all = bind_rows(normal = read_files$null_normal,
#                      exponential = read_files$null_exp,
#                      gamma = read_files$null_gamma,
#                      weibull = read_files$null_weibull,
#                      .id = "distribution")
design_all = readr::read_rds("../hakear-drake/data/design_all.rds")

design_all$design = 
  factor(design_all$design,
         levels = c("null","vary_f", "vary_x", "vary_all"), labels = c("D1", "D2", "D3", "D4"))

design_all$distribution = 
  factor(design_all$distribution,
         levels = c("normal","exponential", "gamma", "weibull"), labels = c("norm", "exp", "gam", "wei"))

library(dplyr)
library(ggplot2)
ggplot2::ggplot(design_all %>% 
                      filter(design == "D1")) +
  ggridges::geom_density_ridges(aes(y = distribution, x = mmpd.x, fill = distribution)) + 
  facet_grid(nx~nfacet) +
  scale_fill_viridis_d() +
  xlab("MMPD")
```

>>>>

<!-- Scagnostics literature -->
This is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which is used to discern meaningful patterns in large collections of scatterplots. Given a set of $v$ variables, there are $v(v-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots. Therefore
for even small $v$, the number of scatterplots can be
large, and scatterplot matrices (SPLOMs) could easily run out of pixels when presenting high-dimensional data. @Dang2014-tw and @wilkinson2005graph provides potential solutions to this, where few characterizations help us to locate anomalies for defining several measures aimed to detect anomalies in density, shape, trend, and
other features in the 2D point scatters.

@@

# Proposed measure: Median Maximum Pairwise Distances (MMPD)

We are interested in assessing structure in probability distributions of the measured variable across bivariate cyclic granularities. We propose a measure called Median Maximum Pairwise Distances (MMPD) to evaluate structure in such a design.

## Idea

@@

# Computation

## Distance between distributions

@@

## Normalize distances

@@

# Simulation study

<!-- ### Distribution of distances -->

<!-- ### Theoretical evidence -->

<!-- JS distances are distributed as chi-squared with $m$ df where we discretize the continuous distribution with $m$ discrete values. Taking sample percentiles to approximate the integral would mean taking $m = 99$. -->
<!-- With large $m$, chi-squared is asymptotically normal by the CLT. Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution. Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made. -->

<!-- ### Empirical evidence -->

<!-- Distribution of JS distances is assumed to be normal but the mean and variance are estimated from the sample, rather than deducing it from the number of discretization used to approximate the continuous distribution. We look at different scenarios, where observations are collected from Normal, Exponential, Chi-squared and Gumbel distribution and found the distribution of JS distances are similar, irrespective of which distribution they are drawn from. -->

<!-- #### Initial distribution of observed variables shown in plot title -->

@@

# Algorithm

@@

@@
